---
title: "ML"
author: 'null'
date: "Saturday, November 15, 2014"
output: html_document
---

###Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).  This background paragraph was taken from the Practical Machine Learning Course Project page by Jeff Leek, PhD, Roger D. Peng, PhD, Brian Caffo, PhD


```{r}
#Read files into R
Train <- read.csv("c:/rWork/pml-training.csv", header=T,
                  stringsAsFactors = FALSE)
Test <- read.csv("c:/rWork/pml-testing.csv", header=T,
                 stringsAsFactors = FALSE)
#Set options to no scientific notations and four digits
options(scipen=999, digits=4)
#Activate caret and rpart libraries
library("caret", lib.loc="~/R/win-library/3.1")
library("rpart", lib.loc="~/R/win-library/3.1")
#Separate Training and Test data
inTrain <- createDataPartition(y=Train$classe, p=.8, list=FALSE)
training <- Train[inTrain,]
trainingTest <- Train[-inTrain,]
dim(training); dim(trainingTest)
#display the classe variable graphically
barchart(training$classe, horiz=TRUE, ylab="Class")
#Fit classe to selected variables


```

###Cross-validation

Cross-validation was used by segmenting the Training dataset into two parts: a smaller training dataset(80%) to be used to build the model and a testing dataset(20%) to be used to validate the model. The goal of cross-validation was to estimate the expected level of fit of the model to a data set that is independent of the data that was used to train the model. It was used to predict the manner in which the exercises were performed as described in the previous paragraph.

###Sample error

Sample error is reduced by increasing the sample size, in this case 80% of the available data. Out-of-sample error was addressed in the same way by providing 20% of the available data when testing the model.


```{r, echo=TRUE}
modFit <- rpart(classe~X+accel_arm_z+cvtd_timestamp+gyros_dumbbell_y+
                    magnet_belt_y+total_accel_forearm, method="class",
                data=trainingTest)
#Predict the fitted rpart object
pred <- predict(modFit, type = "prob")
# display the results
prcp <- printcp(modFit)
# visualize cross-validation results
plcp <- plotcp(modFit)
#detailed summary of splits
summ <- summary(modFit) 
```



